Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-06 Manuscript/TEM_draft.tex
Modified intro+method
Intro: updated paragraph on the why only topology
Method: updated NTS + Tree comparisons + BC explanations

Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case) \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa. This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices. Until now, no attempt has been made to study the impact of this issue on phylogenetic inference using Total Evidence methods.
% NC: THIS SECTION NEEDS WORK!
It is expected that missing in this framework will affect the ability to recover the correct topology as well as the correct branch length estimation \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}. In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies \citep[e.g.][]{manosphylogeny2007,meredithimpacts2011,jetzthe2012}. %TG or at some more specific example instead of the cite: "In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies (e.g. for taxonomical focus \citep{meredithimpacts2011,jetzthe2012}, or for studying evolutionary trends %CITE?\citep[e.g.][]{})"?
Even though branch length is another primordial aspect in phylogenies (namely for timing extinction and/or speciation events \citep[e.g.][]{ronquista2012}), previous studies have empirically assessed the effect of the Total Evidence method on branch length variation \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014} %TG: add:"by using a fixed topology"?
Moreover, variation in branch length are more complex to simulate and to measure when topology varies.
% NC: I don't get the logic of this part 
% because topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e. topological changes are discrete opposed to branch length changes are continuous). 
% NC: Need a better justification of WHY topology is interesting.

%Although correctly inferring branch lengths is also important, we do not consider the inference of branch lengths here.
% NC: WHY??? Difficulties with simulations? Topology is more important? WHY?
%Also, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
% NC: OK this last sentence could be worked in earlier on to explain WHY topology not branch lengths.

@@

\section{Methods}
To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
% NC: I've added numbers in here (125 matrices etc) so it matches the figure legend better. Can you check I've put the right numbers in? I think it may be 124 in some cases...
\begin{enumerate}
\item{Generating the matrix:} \label{step:generate_matrix} \\
We randomly generated a birth-death tree (hereafter called the "true" tree) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix).
\item{Removing data:} \label{step:remove_data} \\
We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$). The resulting 125 matrices are called hereafter "missing-data" matrices.
\item{Inferring phylogenies:} \label{step:build_phylo} \\
We inferred phylogenetic trees from the "complete" matrix and from the 125 "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the "missing-data" trees). Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
\item{Comparing topologies:} \label{step:compare_topo} \\
We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
\end{enumerate}
We repeated these four steps 50 times to account for variation in our random parameters in the simulations.

@@

\subsubsection{Robinson-Foulds distance}
The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details). This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees. This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity. We normalised this metric using the Normalized Tree Similarity index \citep{Bogdanowicz2012}. This index allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details). Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robison-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Noramlised Robison-Foulds distance is lower than zero, the trees are more different than expected by chance.

\subsubsection{Triplets distance}
The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details). This metric measures the position of each taxon and clade in relation to its closest neighbours. It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxon/clade position in the two trees. Therefore this metric sensitive to the conservation of wildcard taxa. We normalised this metric in the same way as we did for the Robison-Foulds distance resulting in the Normalised Triplets distance.

@@

\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion. TG: or no subheading at all?
\label{tree_comparisons}
For the Maximum Likelihood and Bayesian consensus
% NC: is this actually a maximum clade credibility tree - TG: better stick with consensus as described above?
trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance. %For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}). % TG: That is now described in the RF distance subsection just above
%We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree. We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}. TG: explained later on
Also, to take into account the uncertainty of tree inference in both method, instead of using the single Maximum Likelihood and Bayesian consensus trees as above, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees). 
For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above. We repeated this 1000 times for each set of "missing data" trees resulting in 125$\times$1000 values for each metric. We repeated all the paired trees comparisons described above for each of the 50 simulations. We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v. 3.1 \citep{hdrcde}. %Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}. TG: finally we didn't (Andrew's advice - see Battacharyya)

\subsubsection{Summarizing the comparisons}
Finally after doing all theses %stupid
50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework. Additionally, we also assessed the effect of the different methods (i.e. Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).% and the joint effect of the our missing data parameters and the methods.
% TG: the following part might be useless (complete repetition of above)
To assess the ability to recover the "correct" tree topology on our 50 simulations, we used the scores for both normalised metrics (Robinson-Foulds and Triplets distance). We looked at the following effects:
\begin{enumerate}
\item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
\item{The methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
%\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
\end{enumerate}
% TG: end of the repetition

@@

%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
For assessing the effect of our three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$), we calculated the Bhattacharyya Coefficient between each pairwise combination for each parameter state (0\%, 10\%, 25\%, 50\%, 75\%). This resulted in 7875 pairwise distributions pairwise comparisons (triangular $3^5$$\times$$3^5$ matrix). We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}). We repeated this procedure for all the methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}. When two methods display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees, we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods displays dissimilar normalised metrics scores, we expect the Bhattacharyya Coefficients to be close to 0.

@@

\begin{table}[ht]
\label{Tab_Results-Difference_methods}
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
    Bayesian consensus              & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
    Bootstraps                      & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
    bootstraps                      & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
    Boostraps \textit{vs.}          & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
    Bayesian posterior tree         & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}

>>>>

Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case) \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa. This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices. Until now, no attempt has been made to study the impact of this issue on phylogenetic inference using Total Evidence methods.

% NC: THIS SECTION NEEDS WORK!
In this study, we focus on the effect of missing data on our ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}. %CITE-check Lloyd and Friedman or birdy stuff
Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g. \citealt{ronquista2012}), we do not consider branch lengths in this study. This is partially due to difficulties of estimating the effect of missing data simultaneously on branch length and topology, % TG: or "This is partially due to difficulties with simulating branch lengths and topology simultaneously"
but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework. To our knowledge, this question has never been formally assessed.

@@

\section{Methods}
To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
\begin{enumerate}
\item{Generating the matrix:} \label{step:generate_matrix} \\
We randomly generated a birth-death tree (hereafter called the "true" tree) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix).
\item{Removing data:} \label{step:remove_data} \\
We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$). The resulting 125 matrices are called hereafter "missing-data" matrices.
\item{Inferring phylogenies:} \label{step:build_phylo} \\
We inferred phylogenetic trees from the "complete" matrix and from the 125 "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the "missing-data" trees). Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
\item{Comparing topologies:} \label{step:compare_topo} \\
We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
\end{enumerate}
We repeated these four steps 50 times to account for variation in our random parameters in the simulations.

@@

\subsubsection{Robinson-Foulds distance}
The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details). This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees. This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity. We normalised this metric using the Normalized Tree Similarity index \citep{Bogdanowicz2012}. This index allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details). Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robinson-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Normalised Robinson-Foulds distance is lower than zero, the trees are more different than expected by chance.

\subsubsection{Triplets distance}
The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details). This metric measures the position of each taxon and clade in relation to its closest neighbours. It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees. Therefore this metric sensitive to the conservation of wildcard taxa. We normalised this metric in the same way as we did for the Robinson-Foulds distance resulting in the Normalised Triplets distance.

@@

\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion. TG: or no subheading at all? See next section as well.
\label{tree_comparisons}
For the Maximum Likelihood and Bayesian consensus
% NC: is this actually a maximum clade credibility tree - TG: better stick with consensus as described above?
trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance. %For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}). % TG: That is now described in the RF distance subsection just above
%We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree. We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}. TG: explained later on
Also, to take into account the uncertainty of tree inference in both methods cited above, instead of using the single Maximum Likelihood and Bayesian consensus trees, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees). 
For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above. We repeated this 1000 times for each set of "missing data" trees resulting in 125$\times$1000 values for each metric. We repeated all the paired trees comparisons described above for each of the 50 simulations. We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v. 3.1 \citep{hdrcde}. %Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}. TG: finally we didn't (Andrew's advice - see Battacharyya)

\subsubsection{Summarizing the comparisons}
Finally after doing all theses %stupid
50 simulations and comparisons, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework. Additionally, we also assessed the effect of the different methods (i.e. Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).% and the joint effect of the our missing data parameters and the methods.
% TG: the following part might be useless (complete repetition of above)
 To assess the ability to recover the "correct" tree topology on our 50 simulations, we used the scores for both normalised metrics (Robinson-Foulds and Triplets distance). We looked at the following effects:
\begin{enumerate}
\item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
\item{The four "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
%\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
\end{enumerate}
% TG: end of the repetition

@@

%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
For assessing the effect of our three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$), we calculated the Bhattacharyya Coefficient between each pairwise combination for each parameter state (0\%, 10\%, 25\%, 50\%, 75\%). This resulted in 7875 pairwise distributions pairwise comparisons (triangular $3^5$$\times$$3^5$ matrix). We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}). We repeated this procedure for all the "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).

%Explaining the methods comparisons (ANOVA equivalent=among)
Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we compared the similarity between the methods. We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g. Maximum Likelihood\textit{versus} Bayesian consensus). We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e. the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}. When two methods perform similarly in their ability to recover the "correct" tree topology (i.e. they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1. On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.

@@

\begin{table}[ht]
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
    Bayesian consensus              & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
    Bootstraps                      & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
    bootstraps                      & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
    Boostraps \textit{vs.}          & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
    Bayesian posterior tree         & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\label{Tab_Results-Difference_methods}
\end{table}
